---
title: "PLS package comparison"
author: "Lucy Grigoroff"
date: "2023-04-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1.Objective
Compare ropls, metabom8 and RGCCA produced scores and loadings for a PLS analysis. Use the full Cambridge set to make the comparison (ignore the imbalance in control and covid subject numbers). 
#2.Load Data Cambridge LIPO
```{r}
library(fusion)
library(dplyr)
lipo<-local(get(load("~/OneDrive - Murdoch University/datasets/covid19/cambridgeFollowUpPart2/DataElements/covid19_cambridge_ORI_FU1_FU2_PLA_LIPO.daE")))
LIPO<-data.frame(apply(lipo@.Data,2,as.numeric))
LIPO_ANN<-lipo@obsDescr[[1]]

ann<-local(get(load("~/OneDrive - Murdoch University/datasets/covid19/cambridge/DataElements/covid19_cambridge_PLA_ANN_fix.daE")))
#ann<-local(get(load("~/Downloads/covid19_cambridge_PLA_ANN.daE")))
ann = ann@obsDescr[[1]]

#matching the annotation file fo the LIPO_ANN file

ann<-ann[which(ann$sampleID %in% LIPO_ANN$sampleID),]
LIPO<- LIPO[which(LIPO_ANN$sampleID %in% ann$sampleID),]
LIPO_ANN<- LIPO_ANN[which(LIPO_ANN$sampleID %in% ann$sampleID),]

LIPO<- LIPO[match(ann$sampleID, LIPO_ANN$sampleID),]
LIPO_ANN<- LIPO_ANN[match(ann$sampleID, LIPO_ANN$sampleID),]

df<-cbind(LIPO, ann)

#####remove samples with NA for age or group_facet_letter

df2 <-
  df %>%
  filter(!is.na(age)) %>%
  filter(!is.na(group_facet_letter))%>%
  group_by(group_facet_letter, gender) 

#####check for inf values

which(df2 == "Inf", arr.ind = TRUE)

#remove rows with Inf values for lipoprotein quantities
df3 <- df2[-c(49,70,548),]

rm(df, df2)

LipoproteinNames<- as.data.frame(t(t(colnames(df3[, 1:112]))))
LipoproteinNames %>% rename("Lipoprotein"= "V1")
SampleIDs<-as.data.frame(t(t(df3$sampleID)))
SampleIDs %>% rename("sampleID"="V1") 
```

#3.OPLSDA metabom8
plsda is used to fit standard PLS models for classification while splsda performs sparse PLS that embeds feature selection and regularization for the same purpose.
```{r cars}
#pls <- function(X, Y, center = TRUE, scale = 'UV', cv=list(method='k-fold_stratified', k=7, split=2/3), maxPCo = 5, plotting = TRUE)
Metabom8pls<-metabom8::opls(df4[1:112], df4$covid_status, center=TRUE, scale='UV', cv=list(method='MC', k=200, split=2/3), maxPCo=5, plotting=TRUE)
Metabom8scores<-as.data.frame(Metabom8pls@t_pred)
Metabom8loadings<-as.data.frame(t(Metabom8pls@p_pred))

```
#4.PLS RGCCA

If tau is 1 with method set to rgcca, PLS mode A (which favours the summary of each block, aka the outer components). If tau is set to 0 with method rgcca, PLS mode B (which favours the inner model, aka takes into account the relations between the blocks). For RGCCA, to extract the loadings, a plot must be made and the loadings extracted from that. The values of "a" in the model output are the block weight vectors for block j. 
```{r}
library(RGCCA)
#covert gender and covid_status to numeric then as factors

df4<-df3
df4$gender<-as.factor((df4$gender))
df4$covid_status<-as.factor((df4$covid_status))

df4$gender<-as.factor(ifelse(df4$gender=="M",1,0))
df4$covid_status<-as.factor(ifelse(df4$covid_status=="covid",1,0))

Lipop = df4[,1:112] 
metadata= df4[,c("covid_status")]

#set up use for package
blocks <- list(Lipop=Lipop,
               metadata=metadata)

construction<- matrix(c(0,1,
                        1,0), 2,2)

library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
library(caret)

#rgcca model
modCV<- rgcca(blocks = blocks,
               method= "rgcca",
               connection = construction,
               tau = rep(1, length(blocks)), # has to be one for sgcca, 0 for rgcca (PLS mode B) 1 for rgcca (PLS mode A)
               sparsity = c(1, 1),
               response = length(blocks),
               ncomp = rep(1, length(blocks)),
               scheme = "centroid", # factorial and centroid allows for negative correlation
               verbose = TRUE)

RGCCAscores<-as.data.frame(modCV[["Y"]][["Lipop"]])
RGCCAloadings<-as.data.frame(modCV[["a"]][["Lipop"]])#gives information for variable selection
 # RGCCALoad<-plot(modCV, type = "loadings", n_mark=113)
 # RGCCAloadings<-cbind(as.data.frame(RGCCALoad[["data"]][["x"]]), as.data.frame(RGCCALoad[["data"]][["y"]])) %>%
 #   rename("V1"=1, "Lipoprotein"=2)
 # RGCCAloadings<-RGCCAloadings[-1,]


#prep model output for graphing
# LipopScore<-modCV[["Y"]][["Lipop"]]
# MetaScore<-modCV[["Y"]][["metadata"]]
# df<-cbind((LipopScore), (MetaScore))
# df<-as.data.frame(df)
# df<-df %>%
#   rename( "LipopScore"=1, "MetaScore"=2)
# 
# #graph and correlation 
# ggplot(data = df, aes(x = MetaScore, y = LipopScore, color=df4$gender, shape=df4$covid_status)) +
#         geom_point() +
#         labs(title = "SGCCA Produced Scores",
#              x = "Metadata Scores",
#              y = "Lipoprotein Scores")+
#    geom_smooth(method = "lm", se = FALSE)
# 
# 
# cor.test(df$MetaScore, df$LipopScore, method="pearson") ##sample estimates:cor -0.7319135, 95% CI: (-0.8154106, -0.6185827)
# 
# RGCCAloadings<-modCV[["a"]][["Lipop"]]#gives information for variable selection
```

#5.OPLSDA ropls
## Including Plots
scoreMN Numerical matrix of x scores (T; dimensions: nrow(x) x predI) X = TP' + E; Y = TC' + F

loadingMN Numerical matrix of x loadings (P; dimensions: ncol(x) x predI) X = TP' + E

```{r pressure, echo=FALSE}

roplsresult<-ropls::opls(df3[,1:112], df3$covid_status, predI = NA, orthoI = 0, algoC='nipals', crossvalI = 7)
#roplsres<-ropls::opls(df3[,1:112], df3$covid_status)
roplsscores<-as.data.frame(roplsresult@scoreMN)
roplsloadings<-as.data.frame(roplsresult@loadingMN)


```

#6.Predictive loadings comparison
```{r}
PLSload<- as.data.frame((cbind(LipoproteinNames, roplsloadings$p1, Metabom8loadings$V1, RGCCAloadings$V1))) %>%
  rename("ropls" ="roplsloadings$p1","metabom8" ="Metabom8loadings$V1", "rgcca"="RGCCAloadings$V1")
 

library(reshape2)
df <- melt(PLSload ,  id.vars = 'V1', variable.name = 'Method')

# plot on same grid, each series colored differently -- 
# good if the series have same scale
ggplot(df, aes(V1,value)) + 
  geom_point(aes(colour = Method)) +
  geom_line() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(x="Lipoprotein", y="Loadings")

 
#plots for correlation 
ggplot(data=PLSload, aes(ropls, metabom8)) + geom_point()+stat_summary(fun.data=mean_cl_normal) + 
  geom_smooth(method='lm', formula= y~x)+ theme_minimal()+ggtitle("Loadings")
cor.test(PLSload$ropls, PLSload$metabom8, method="pearson")

ggplot(data=PLSload, aes(ropls, rgcca)) + geom_point()+stat_summary(fun.data=mean_cl_normal) + 
  geom_smooth(method='lm', formula= y~x)+ theme_minimal() + ggtitle("Loadings")
cor.test(PLSload$ropls, PLSload$rgcca, method="pearson")

ggplot(data=PLSload, aes(metabom8, rgcca)) + geom_point()+stat_summary(fun.data=mean_cl_normal) + 
  geom_smooth(method='lm', formula= y~x)+ theme_minimal()+ggtitle("Loadings")
cor.test(PLSload$metabom8, PLSload$rgcca, method="pearson")


```
#7.Predictive scores comparison
```{r}
PLSscore<- as.data.frame((cbind(SampleIDs, roplsscores$p1, Metabom8scores$V1, RGCCAscores$comp1))) %>%
  rename("ropls" ="roplsscores$p1","metabom8" ="Metabom8scores$V1", "rgcca"="RGCCAscores$comp1")
 

library(reshape2)
df <- melt(PLSscore ,  id.vars = 'V1', variable.name = 'Method')

# plot on same grid, each series colored differently -- 
# good if the series have same scale
ggplot(df, aes(V1,value)) + 
  geom_point(aes(colour = Method)) +
  #geom_line() +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(axis.text.x = element_blank())+
  labs(x="SampleID", y="Scores")


#plots for correlation 
ggplot(data=PLSscore, aes(ropls, metabom8)) + geom_point()+stat_summary(fun.data=mean_cl_normal) + 
  geom_smooth(method='lm', formula= y~x)+ theme_minimal()+ggtitle("Scores")
cor.test(PLSscore$ropls, PLSscore$metabom8, method="pearson")

######has exact correlation of 1 between rgcca and ropls, but not rgcca and metabom8. One is is based off the other (rgcca and ropls) at best guess. Can rescale and revisit if tau should be 0 or 1 for rgcca. One of which will correspond to opls and the other pls. 
ggplot(data=PLSscore, aes(ropls, rgcca)) + geom_point()+stat_summary(fun.data=mean_cl_normal) + 
  geom_smooth(method='lm', formula= y~x)+ theme_minimal()+ggtitle("Scores")
cor.test(PLSscore$ropls, PLSscore$rgcca, method="pearson")

ggplot(data=PLSscore, aes(metabom8, rgcca)) + geom_point()+stat_summary(fun.data=mean_cl_normal) + 
  geom_smooth(method='lm', formula= y~x) + theme_minimal()+ggtitle("Scores")
cor.test(PLSscore$metabom8, PLSscore$rgcca, method="pearson")


```

#8.ropls correlation

The ropls package won't perform an OPLS(DA) on teh large cambridge dataset. Below is a comparison of the predictive scores and loadings using a smaller, evenly grouped dataset (same number of covid/controls and equal gender rep in each group). 

The predictive scores and loadings are highly correlated and so the use of the PLSDA on the large cambridge data set for predicitve scores and loadings (calculated above in section 5. and compared in sections 6. and 7. ) is justifiable. 
```{r}

set.seed(1)

malecovid<-df3 %>%
  filter(covid_status=="covid" ,time_point==0, gender=="M")
femalecovid<-df3 %>%
  filter(covid_status=="covid" ,time_point==0, gender=="F")
controls<-df3 %>%
  filter(covid_status=="control", time_point==0)

x<-femalecovid[sample(nrow(femalecovid), 25), ]
y<-malecovid[sample(nrow(malecovid), 20), ]

df5<-rbind(controls, x, y)
df5$sampleID
df5Samples<-cat(paste((df5$sampleID), type=","))

#new ropls for PLS-DA with correct control/covid and gender balance
roplsPLSDA<-ropls::opls(df5[,1:112], df5$covid_status, predI = NA, orthoI = 0, algoC='nipals', crossvalI = 7)#typeC is PLS-DA
roplsPLSDAscores<-as.data.frame(roplsPLSDA@scoreMN)
roplsPLSDAloadings<-as.data.frame(roplsPLSDA@loadingMN)

XscaledPLSDA<-roplsPLSDA@suppLs[["xModelMN"]]

corPLSDA<-cor((roplsPLSDAscores$p1),(XscaledPLSDA))
corPLSDA_df<-data.frame(t(corPLSDA))
corPLSDA_df2<-cbind(colnames(df5[,1:112]),corPLSDA_df)
colnames(corPLSDA_df2) <- c("Lipoprotein", "PLSDA_correlation")
PLSDAsig <- corPLSDA_df2 %>%
  filter(PLSDA_correlation<(-0.5)|PLSDA_correlation>0.5)



#new ropls for OPLS-DA
roplsOPLSDA<-ropls::opls(df5[,1:112], df5$covid_status, predI = 1, orthoI = NA, algoC='nipals', crossvalI = 7)#typeC is OPLS-DA
roplsOPLSDAscores<-as.data.frame(roplsOPLSDA@scoreMN)
roplsOPLSDAloadings<-as.data.frame(roplsOPLSDA@loadingMN)

XscaledOPLSDA<-roplsOPLSDA@suppLs[["xModelMN"]]

corOPLSDA<-cor((roplsOPLSDAscores$p1),(XscaledOPLSDA))
corOPLSDA_df<-data.frame(t(corOPLSDA))
corOPLSDA_df2<-cbind(colnames(df5[,1:112]),corOPLSDA_df)
colnames(corOPLSDA_df2) <- c("Lipoprotein", "OPLSDA_correlation")
OPLSDAsig <- corOPLSDA_df2 %>%
  filter(OPLSDA_correlation<(-0.5)|OPLSDA_correlation>0.5)
#compare pls-da and opls-da loadings and scores since opls-da doesn't work with the full set in df3, only pls-da does
proof<-as.data.frame((cbind(roplsOPLSDAscores$p1, roplsPLSDAscores$p1))) 


ggplot(data=proof, aes(roplsOPLSDAscores$p1, roplsPLSDAscores$p1)) + geom_point()+stat_summary(fun.data=mean_cl_normal) + 
  geom_smooth(method='lm', formula= y~x) + theme_minimal()+ggtitle("ropls Scores") + labs(x="oplsda", y="plsda")
cor.test(roplsOPLSDAscores$p1, roplsPLSDAscores$p1, method="pearson")

proof<-as.data.frame((cbind(roplsOPLSDAloadings$p1, roplsPLSDAloadings$p1)))
ggplot(data=proof, aes(roplsOPLSDAloadings$p1, roplsPLSDAloadings$p1)) + geom_point()+stat_summary(fun.data=mean_cl_normal) + 
  geom_smooth(method='lm', formula= y~x) + theme_minimal()+ggtitle("ropls Loadings") + labs(x="oplsda", y="plsda")
cor.test(roplsOPLSDAloadings$p1, roplsPLSDAloadings$p1, method="pearson")
#comparison of PLS-DA, OPLS-DA and Univariate p-values 

comp<-full_join(UnivariateSignificant,PLSDAsig,by=c("Lipoprotein"))
comp<-full_join(comp, OPLSDAsig,by=c("Lipoprotein"))
# test<-full_join(test, sigRGCCA, by=c("Lipoprotein"))


V=list(UV=UnivariateSignificant$Lipoprotein, PLSDA=PLSDAsig$Lipoprotein, OPLSDA=OPLSDAsig$Lipoprotein)
ggVennDiagram(V, color = "black", lwd = 0.8, lty = 1) + 
  scale_fill_gradient(low = "#F4FAFE", high = "#4981BF")
```

#9. Orthoginal
Only metabom8 and ropls calculated and compared, RGCCA does not provide OPLS/orthoginal values.
```{r}
#metabom8
Metabom8pls<-metabom8::opls(df5[1:112], df5$covid_status, center=TRUE, scale='UV', cv=list(method='MC', k=200, split=2/3), maxPCo=5, plotting=TRUE)
Metabom8Orthscores<-as.data.frame(Metabom8pls@t_orth)
Metabom8Orthloadings<-as.data.frame(t(Metabom8pls@p_orth))

#ropls
roplsOPLSDA<-ropls::opls(df5[,1:112], df5$covid_status, predI = 1, orthoI = NA, algoC='nipals', crossvalI = 7)#typeC is OPLS-DA
roplsOPLSDAOrthscores<-as.data.frame(roplsOPLSDA@orthoScoreMN)
roplsOPLSDAOrthloadings<-as.data.frame(roplsOPLSDA@orthoLoadingMN)

#comparative plots
##Scores
ORTHscore<- as.data.frame((cbind(df5$sampleID, roplsOPLSDAOrthscores$o1, Metabom8Orthscores$V1))) %>%
  rename(,"ropls" ="V2","metabom8" ="V3", )
ORTHscore$ropls<-as.numeric(ORTHscore$ropls) 
ORTHscore$metabom8<-as.numeric(ORTHscore$metabom8) 

library(reshape2)
df <- melt(ORTHscore ,  id.vars = 'V1', variable.name = 'Method')

### plot on same grid, each series colored differently -- 
### good if the series have same scale
ggplot(df, aes(V1,value)) + 
  geom_point(aes(colour = Method)) +
  #geom_line() +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(axis.text.x = element_blank())+
  labs(x="SampleID", y="Scores")


###plots for correlation 
ggplot(data=ORTHscore, aes(ropls, metabom8)) + geom_point()+stat_summary(fun.data=mean_cl_normal) + 
  geom_smooth(method='lm', formula= y~x)+ theme_minimal()+ggtitle("Orthogonal Scores")
cor.test(ORTHscore$ropls, ORTHscore$metabom8, method="pearson")


##Loadings
ORTHloading<- as.data.frame((cbind(LipoproteinNames, roplsOPLSDAloadings$p1, Metabom8Orthloadings$V1))) %>%
  rename("ropls" ="roplsOPLSDAloadings$p1","metabom8" ="Metabom8Orthloadings$V1", )
 

library(reshape2)
df <- melt(ORTHloading ,  id.vars = 'V1', variable.name = 'Method')

### plot on same grid, each series colored differently -- 
### good if the series have same scale
ggplot(df, aes(V1,value)) + 
  geom_point(aes(colour = Method)) +
  #geom_line() +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(axis.text.x = element_blank())+
  labs(x="SampleID", y="Scores")


###plots for correlation 
ggplot(data=ORTHloading, aes(ropls, metabom8)) + geom_point()+stat_summary(fun.data=mean_cl_normal) + 
  geom_smooth(method='lm', formula= y~x)+ theme_minimal()+ggtitle("Orthogonal Loadings")
cor.test(ORTHloading$ropls, ORTHloading$metabom8, method="pearson")
```

#Load Data Biogune Amino Acids
```{r}
ann<-local(get(load("~/OneDrive - Murdoch University/datasets/covid19/bioGune/dataElements/covid19_bioGune_PLA_ANN.daE")))
aaMeta1<-ann@obsDescr[[1]]

aa<-local(get(load(file.path("~/OneDrive - Murdoch University/datasets/covid19/bioGune/dataElements/covid19_bioGune_PLA_AA.daE"))))
aaData<-data.frame(apply(aa@.Data,2,as.numeric)) #Amino Acids
aaMeta2<-aa@obsDescr[[1]] #metadata from covid19_mauritius_PLA_LIPO.daE file

# match Amino Acids and both Annotations
aaMeta1<-aaMeta1[which(aaMeta1$sampleID %in% aaMeta2$sampleID),]
aaData<-aaData[which(aaMeta2$sampleID %in% aaMeta1$sampleID),]
aaMeta2<-aaMeta2[which(aaMeta2$sampleID %in% aaMeta1$sampleID),]

aaMetaF<-merge(aaMeta1,
                  aaMeta2,
                  by.x ="sampleID",by.y = "sampleID")
rm(aaMeta1, aaMeta2, aa, ann)
aaData<-aaData[,-grep("13C|AccQTag",colnames(aaData))]
aaData<-aaData[,-which(colnames(aaData) %in% c("beta.aminoisobutyric.acid","beta.alanine","gamma.aminobutyric.acid","Cystine", "Sarcosine"))]

#remove X3-methylhistidine (too many NAs)
idx<-which(is.na(aaData),arr.ind = TRUE) #see entries with NA
col_idx<-as.numeric(unique(idx[,2])) #identify the coloumn numbers
aaData<- aaData[,-2]#remove the column

aaData$aa_na_count <- apply(aaData, 1, function(x) sum(is.na(x))) #make column of how many NAs exist in each row

BIOaa<-cbind(aaData, aaMetaF)
rm(aaData, aaMetaF)

idx<-which(BIOaa$aa_na_count>0)
BIOaa<-BIOaa[-idx,]

which(is.na(BIOaa$sex))
unique(BIOaa$sex)

unique(BIOaa$group)
BIOaa$covid_status<-ifelse(BIOaa$group =="COVID-pos","covid", ifelse(BIOaa$group=="preCOVID","control", "control"))

table(BIOaa$sex, BIOaa$covid_status)

#smaller dataset
set.seed(1)

malecovid<-BIOaa %>%
  filter(covid_status=="covid" ,sex=="male")
femalecovid<-BIOaa %>%
  filter(covid_status=="covid" ,sex=="female")
controls<-BIOaa %>%
  filter(covid_status=="control")

x<-femalecovid[sample(nrow(femalecovid), 60), ]
y<-malecovid[sample(nrow(malecovid), 60), ]

BIOaa_small<-rbind(controls, x, y)


rm(x, y, controls, femalecovid, malecovid)

```
#OPLS Metabom8
```{r}
Metabom8pls<-metabom8::opls(BIOaa[1:23], BIOaa$covid_status, center=TRUE, scale='UV', cv=list(method='MC', k=200, split=2/3), maxPCo=0, plotting=TRUE)
Metabom8scores<-as.data.frame(Metabom8pls@t_pred)
Metabom8loadings<-as.data.frame(t(Metabom8pls@p_pred))
```

#OPLS ropls

```{r}

roplsresult<-ropls::opls(BIOaa[,1:23], BIOaa$covid_status, predI = 1, orthoI = NA, algoC='nipals', crossvalI = 7)
#roplsres<-ropls::opls(df3[,1:112], df3$covid_status)
roplsscores<-as.data.frame(roplsresult@scoreMN)
roplsloadings<-as.data.frame(roplsresult@loadingMN)
```

#loadings comparison
```{r}
AminoAcids<-as.data.frame(colnames(BIOaa_small))

PLSload<- as.data.frame((cbind(AminoAcids[1:23,], roplsloadings$p1, Metabom8loadings$V1))) %>%
  rename("AminoAcids"="V1","ropls" ="V3","metabom8" ="V2")
PLSload$metabom8<-as.numeric(PLSload$metabom8)
PLSload$ropls<-as.numeric(PLSload$ropls)
library(reshape2)
df <- melt(PLSload ,  id.vars = 'AminoAcids', variable.name = 'Method')

# plot on same grid, each series colored differently -- 
# good if the series have same scale
ggplot(df, aes(AminoAcids,value)) + 
  geom_point(aes(colour = Method)) +
  geom_line() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(x="AminoAcids", y="Loadings")

 
#plots for correlation 
ggplot(data=PLSload, aes(ropls, metabom8)) + geom_point()+ stat_summary(fun.data=mean_cl_normal) + geom_smooth(method='lm', formula= y~x) + theme_minimal() + ggtitle("Loadings")
  cor.test(PLSload$metabom8, PLSload$ropls, method="pearson")

```

#Scores comparison
```{r}
SampleID<-as.data.frame(BIOaa$sampleID)
PLSscore<- as.data.frame((cbind(SampleID, roplsscores$p1, Metabom8scores$V1))) %>%
  rename("SampleID"="BIOaa$sampleID","ropls" ="roplsscores$p1","metabom8" ="Metabom8scores$V1")
 

library(reshape2)
df <- melt(PLSscore ,  id.vars = 'SampleID', variable.name = 'Method')

# plot on same grid, each series colored differently -- 
# good if the series have same scale
ggplot(df, aes(SampleID,value)) + 
  geom_point(aes(colour = Method)) +
  #geom_line() +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(axis.text.x = element_blank())+
  labs(x="SampleID", y="Scores")


#plots for correlation 
ggplot(data=PLSscore, aes(ropls, metabom8)) + geom_point()+stat_summary(fun.data=mean_cl_normal) + 
  geom_smooth(method='lm', formula= y~x)+ theme_minimal()+ggtitle("Scores")
cor.test(PLSscore$ropls, PLSscore$metabom8, method="pearson")
```
#PLS RGCCA

If tau is 1 with method set to rgcca, PLS mode A (which favours the summary of each block, aka the outer components). If tau is set to 0 with method rgcca, PLS mode B (which favours the inner model, aka takes into account the relations between the blocks). For RGCCA, to extract the loadings, a plot must be made and the loadings extracted from that. The values of "a" in the model output are the block weight vectors for block j. 
```{r}
library(RGCCA)
#covert gender and covid_status to numeric then as factors

BIOaaRGCCA<-BIOaa
BIOaaRGCCA$sex<-as.factor((BIOaaRGCCA$sex))
BIOaaRGCCA$covid_status<-as.factor((BIOaaRGCCA$covid_status))

BIOaaRGCCA$sex<-as.factor(ifelse(BIOaaRGCCA$sex=="M",1,0))
BIOaaRGCCA$covid_status<-as.factor(ifelse(BIOaaRGCCA$covid_status=="covid",1,0))

AminoAcids = BIOaaRGCCA[,1:23] 
metadata= BIOaaRGCCA[,c("covid_status")]

#set up use for package
blocks <- list(AminoAcids=AminoAcids,
               metadata=metadata)

construction<- matrix(c(0,1,
                        1,0), 2,2)

library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
library(caret)

#rgcca model
modCV<- rgcca(blocks = blocks,
               method= "rgcca",
               connection = construction,
               tau = rep(1, length(blocks)), # has to be one for sgcca, 0 for rgcca (PLS mode B) 1 for rgcca (PLS mode A)
               sparsity = c(1, 1),
               response = length(blocks),
               ncomp = rep(1, length(blocks)),
               scheme = "centroid", # factorial and centroid allows for negative correlation
               verbose = TRUE)

RGCCAscores<-as.data.frame(modCV[["Y"]][["AminoAcids"]])
RGCCAloadings<-as.data.frame(modCV[["a"]][["AminoAcids"]])#gives information for variable selection
```

#PLS ropls

```{r}
roplsresult<-ropls::opls(BIOaa[,1:23], BIOaa$covid_status, predI = NA, orthoI = 0, algoC='nipals', crossvalI = 7)
#roplsres<-ropls::opls(df3[,1:112], df3$covid_status)
roplsscores<-as.data.frame(roplsresult@scoreMN)
roplsloadings<-as.data.frame(roplsresult@loadingMN)
```

#loadings comparison
```{r}
AminoAcids<-as.data.frame(colnames(BIOaa_small))

PLSload<- as.data.frame((cbind(AminoAcids[1:23,], roplsloadings$p1, RGCCAloadings$V1))) %>%
  rename("AminoAcids"="V1","ropls" ="V3","rgcca" ="V2")
PLSload$rgcca<-as.numeric(PLSload$rgcca)
PLSload$ropls<-as.numeric(PLSload$ropls)
library(reshape2)
df <- melt(PLSload ,  id.vars = 'AminoAcids', variable.name = 'Method')

# plot on same grid, each series colored differently -- 
# good if the series have same scale
ggplot(df, aes(AminoAcids,value)) + 
  geom_point(aes(colour = Method)) +
  geom_line() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(x="AminoAcids", y="Loadings")

 
#plots for correlation 
ggplot(data=PLSload, aes(ropls, rgcca)) + geom_point()+ stat_summary(fun.data=mean_cl_normal) + geom_smooth(method='lm', formula= y~x) + theme_minimal() + ggtitle("Loadings")
  cor.test(PLSload$metabom8, PLSload$ropls, method="pearson")

```

#Scores comparison
```{r}
SampleID<-as.data.frame(BIOaa$sampleID)
PLSscore<- as.data.frame((cbind(SampleID, roplsscores$p1, RGCCAscores$comp1))) %>%
  rename("SampleID"="BIOaa$sampleID","ropls" ="roplsscores$p1","rgcca" ="RGCCAscores$comp1")
 

library(reshape2)
df <- melt(PLSscore ,  id.vars = 'SampleID', variable.name = 'Method')

# plot on same grid, each series colored differently -- 
# good if the series have same scale
ggplot(df, aes(SampleID,value)) + 
  geom_point(aes(colour = Method)) +
  #geom_line() +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(axis.text.x = element_blank())+
  labs(x="SampleID", y="Scores")


#plots for correlation 
ggplot(data=PLSscore, aes(ropls, rgcca)) + geom_point()+stat_summary(fun.data=mean_cl_normal) + 
  geom_smooth(method='lm', formula= y~x)+ theme_minimal()+ggtitle("Scores")
cor.test(PLSscore$ropls, PLSscore$rgcca, method="pearson")
```