---
title: "PLS package comparison"
author: "Lucy Grigoroff"
date: "2023-04-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Objective
Compare ropls, metabom8 and RGCCA produced scores and loadings for a PLS analysis. Use the full Cambridge set to make the comparison (ignore the imbalance in control and covid subject numbers). 
#load data
```{r}
library(fusion)
library(dplyr)
lipo<-local(get(load("~/OneDrive - Murdoch University/datasets/covid19/cambridgeFollowUpPart2/DataElements/covid19_cambridge_ORI_FU1_FU2_PLA_LIPO.daE")))
LIPO<-data.frame(apply(lipo@.Data,2,as.numeric))
LIPO_ANN<-lipo@obsDescr[[1]]

ann<-local(get(load("~/OneDrive - Murdoch University/datasets/covid19/cambridge/DataElements/covid19_cambridge_PLA_ANN_fix.daE")))
#ann<-local(get(load("~/Downloads/covid19_cambridge_PLA_ANN.daE")))
ann = ann@obsDescr[[1]]

#matching the annotation file fo the LIPO_ANN file

ann<-ann[which(ann$sampleID %in% LIPO_ANN$sampleID),]
LIPO<- LIPO[which(LIPO_ANN$sampleID %in% ann$sampleID),]
LIPO_ANN<- LIPO_ANN[which(LIPO_ANN$sampleID %in% ann$sampleID),]

LIPO<- LIPO[match(ann$sampleID, LIPO_ANN$sampleID),]
LIPO_ANN<- LIPO_ANN[match(ann$sampleID, LIPO_ANN$sampleID),]

df<-cbind(LIPO, ann)

#####remove samples with NA for age or group_facet_letter

df2 <-
  df %>%
  filter(!is.na(age)) %>%
  filter(!is.na(group_facet_letter))%>%
  group_by(group_facet_letter, gender) 

#####check for inf values

which(df2 == "Inf", arr.ind = TRUE)

#remove rows with Inf values for lipoprotein quantities
df3 <- df2[-c(49,70,548),]

rm(df, df2)

LipoproteinNames<- as.data.frame(t(t(colnames(df3[, 1:112]))))
LipoproteinNames %>% rename("Lipoprotein"= "V1")
SampleIDs<-as.data.frame(t(t(df3$sampleID)))
SampleIDs %>% rename("sampleID"="V1") 
```

#metabom8
plsda is used to fit standard PLS models for classification while splsda performs sparse PLS that embeds feature selection and regularization for the same purpose.
```{r cars}
#pls <- function(X, Y, center = TRUE, scale = 'UV', cv=list(method='k-fold_stratified', k=7, split=2/3), maxPCo = 5, plotting = TRUE)
Metabom8pls<-metabom8::opls(df3[1:112], df3$covid_status, center=TRUE, scale='UV', cv=list(method='MC', k=200, split=2/3), maxPCo=5, plotting=TRUE)
Metabom8scores<-as.data.frame(Metabom8pls@t_pred)
Metabom8loadings<-as.data.frame(t(Metabom8pls@p_pred))

```
#RGCCA

If tau is 1 with method set to rgcca, PLS mode A (which favours the summary of each block, aka the outer components). If tau is set to 0 with method rgcca, PLS mode B (which favours the inner model, aka takes into account the relations between the blocks). For RGCCA, to extract the loadings, a plot must be made and the loadings extracted from that. The values of "a" in the model output are the block weight vectors for block j. 
```{r}
library(RGCCA)
#covert gender and covid_status to numeric then as factors

df4<-df3
df4$gender<-as.factor((df4$gender))
df4$covid_status<-as.factor((df4$covid_status))

df4$gender<-as.factor(ifelse(df4$gender=="M",1,0))
df4$covid_status<-as.factor(ifelse(df4$covid_status=="covid",1,0))

Lipop = df4[,1:112] 
metadata= df4[,c("covid_status")]

#set up use for package
blocks <- list(Lipop=Lipop,
               metadata=metadata)

construction<- matrix(c(0,1,
                        1,0), 2,2)

library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
library(caret)

#rgcca model
modCV<- rgcca(blocks = blocks,
               method= "rgcca",
               connection = construction,
               tau = rep(1, length(blocks)), # has to be one for sgcca, 0 for rgcca (PLS mode B) 1 for rgcca (PLS mode A)
               sparsity = c(1, 1),
               response = length(blocks),
               ncomp = rep(1, length(blocks)),
               scheme = "centroid", # factorial and centroid allows for negative correlation
               verbose = TRUE)

RGCCAscores<-as.data.frame(modCV[["Y"]][["Lipop"]])
RGCCAloadings<-as.data.frame(modCV[["a"]][["Lipop"]])#gives information for variable selection
 # RGCCALoad<-plot(modCV, type = "loadings", n_mark=113)
 # RGCCAloadings<-cbind(as.data.frame(RGCCALoad[["data"]][["x"]]), as.data.frame(RGCCALoad[["data"]][["y"]])) %>%
 #   rename("V1"=1, "Lipoprotein"=2)
 # RGCCAloadings<-RGCCAloadings[-1,]


#prep model output for graphing
# LipopScore<-modCV[["Y"]][["Lipop"]]
# MetaScore<-modCV[["Y"]][["metadata"]]
# df<-cbind((LipopScore), (MetaScore))
# df<-as.data.frame(df)
# df<-df %>%
#   rename( "LipopScore"=1, "MetaScore"=2)
# 
# #graph and correlation 
# ggplot(data = df, aes(x = MetaScore, y = LipopScore, color=df4$gender, shape=df4$covid_status)) +
#         geom_point() +
#         labs(title = "SGCCA Produced Scores",
#              x = "Metadata Scores",
#              y = "Lipoprotein Scores")+
#    geom_smooth(method = "lm", se = FALSE)
# 
# 
# cor.test(df$MetaScore, df$LipopScore, method="pearson") ##sample estimates:cor -0.7319135, 95% CI: (-0.8154106, -0.6185827)
# 
# RGCCAloadings<-modCV[["a"]][["Lipop"]]#gives information for variable selection
```

#ropls
## Including Plots
scoreMN Numerical matrix of x scores (T; dimensions: nrow(x) x predI) X = TP' + E; Y = TC' + F

loadingMN Numerical matrix of x loadings (P; dimensions: ncol(x) x predI) X = TP' + E

```{r pressure, echo=FALSE}

roplsresult<-ropls::opls(df3[,1:112], df3$covid_status, predI = 1, orthoI = NA, algoC='nipals', crossvalI = 7)
#roplsres<-ropls::opls(df3[,1:112], df3$covid_status)
roplsscores<-as.data.frame(roplsresult@scoreMN)
roplsloadings<-as.data.frame(roplsresult@loadingMN)
```

#loadings comparison
```{r}
PLSload<- as.data.frame((cbind(LipoproteinNames, roplsloadings$p1, Metabom8loadings$V1, RGCCAloadings$V1))) %>%
  rename("ropls" ="roplsloadings$p1","metabom8" ="Metabom8loadings$V1", "rgcca"="RGCCAloadings$V1")
 

library(reshape2)
df <- melt(PLSload ,  id.vars = 'V1', variable.name = 'Method')

# plot on same grid, each series colored differently -- 
# good if the series have same scale
ggplot(df, aes(V1,value)) + 
  geom_point(aes(colour = Method)) +
  geom_line() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(x="Lipoprotein", y="Loadings")

 
#plots for correlation 
ggplot(data=PLSload, aes(ropls, metabom8)) + geom_point()+stat_summary(fun.data=mean_cl_normal) + 
  geom_smooth(method='lm', formula= y~x)+ theme_minimal()+ggtitle("Loadings")
cor.test(PLSload$ropls, PLSload$metabom8, method="pearson")

ggplot(data=PLSload, aes(ropls, rgcca)) + geom_point()+stat_summary(fun.data=mean_cl_normal) + 
  geom_smooth(method='lm', formula= y~x)+ theme_minimal() + ggtitle("Loadings")
cor.test(PLSload$ropls, PLSload$rgcca, method="pearson")

ggplot(data=PLSload, aes(metabom8, rgcca)) + geom_point()+stat_summary(fun.data=mean_cl_normal) + 
  geom_smooth(method='lm', formula= y~x)+ theme_minimal()+ggtitle("Loadings")
cor.test(PLSload$metabom8, PLSload$rgcca, method="pearson")


```
#scores comparison
```{r}
PLSscore<- as.data.frame((cbind(SampleIDs, roplsscores$p1, Metabom8scores$V1, RGCCAscores$comp1))) %>%
  rename("ropls" ="roplsscores$p1","metabom8" ="Metabom8scores$V1", "rgcca"="RGCCAscores$comp1")
 

library(reshape2)
df <- melt(PLSscore ,  id.vars = 'V1', variable.name = 'Method')

# plot on same grid, each series colored differently -- 
# good if the series have same scale
ggplot(df, aes(V1,value)) + 
  geom_point(aes(colour = Method)) +
  #geom_line() +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(axis.text.x = element_blank())+
  labs(x="SampleID", y="Scores")


#plots for correlation 
ggplot(data=PLSscore, aes(ropls, metabom8)) + geom_point()+stat_summary(fun.data=mean_cl_normal) + 
  geom_smooth(method='lm', formula= y~x)+ theme_minimal()+ggtitle("Scores")
cor.test(PLSscore$ropls, PLSscore$metabom8, method="pearson")

######has exact correlation of 1 between rgcca and ropls, but not rgcca and metabom8. One is is based off the other (rgcca and ropls) at best guess. Can rescale and revisit if tau should be 0 or 1 for rgcca. One of which will correspond to opls and the other pls. 
ggplot(data=PLSscore, aes(ropls, rgcca)) + geom_point()+stat_summary(fun.data=mean_cl_normal) + 
  geom_smooth(method='lm', formula= y~x)+ theme_minimal()+ggtitle("Scores")
cor.test(PLSscore$ropls, PLSscore$rgcca, method="pearson")

ggplot(data=PLSscore, aes(metabom8, rgcca)) + geom_point()+stat_summary(fun.data=mean_cl_normal) + 
  geom_smooth(method='lm', formula= y~x) + theme_minimal()+ggtitle("Scores")
cor.test(PLSscore$metabom8, PLSscore$rgcca, method="pearson")


```
#ropls correlation
```{r}

set.seed(1)

malecovid<-df3 %>%
  filter(covid_status=="covid" ,time_point==0, gender=="M")
femalecovid<-df3 %>%
  filter(covid_status=="covid" ,time_point==0, gender=="F")
controls<-df3 %>%
  filter(covid_status=="control", time_point==0)

x<-femalecovid[sample(nrow(femalecovid), 25), ]
y<-malecovid[sample(nrow(malecovid), 20), ]

df5<-rbind(controls, x, y)
df5$sampleID
df5Samples<-cat(paste((df5$sampleID), type=","))

#new ropls for PLS-DA with correct control/covid and gender balance
roplsPLSDA<-ropls::opls(df5[,1:112], df5$covid_status, predI = NA, orthoI = 0, algoC='nipals', crossvalI = 7)#typeC is PLS-DA
roplsPLSDAscores<-as.data.frame(roplsPLSDA@scoreMN)
roplsPLSDAloadings<-as.data.frame(roplsPLSDA@loadingMN)

XscaledPLSDA<-roplsPLSDA@suppLs[["xModelMN"]]

corPLSDA<-cor((roplsPLSDAscores$p1),(XscaledPLSDA))
corPLSDA_df<-data.frame(t(corPLSDA))
corPLSDA_df2<-cbind(colnames(df5[,1:112]),corPLSDA_df)
colnames(corPLSDA_df2) <- c("Lipoprotein", "PLSDA_correlation")
PLSDAsig <- corPLSDA_df2 %>%
  filter(PLSDA_correlation<(-0.5)|PLSDA_correlation>0.5)



#new ropls for OPLS-DA
roplsOPLSDA<-ropls::opls(df5[,1:112], df5$covid_status, predI = 1, orthoI = NA, algoC='nipals', crossvalI = 7)#typeC is OPLS-DA
roplsOPLSDAscores<-as.data.frame(roplsOPLSDA@scoreMN)
roplsOPLSDAloadings<-as.data.frame(roplsOPLSDA@loadingMN)

XscaledOPLSDA<-roplsOPLSDA@suppLs[["xModelMN"]]

corOPLSDA<-cor((roplsOPLSDAscores$p1),(XscaledOPLSDA))
corOPLSDA_df<-data.frame(t(corOPLSDA))
corOPLSDA_df2<-cbind(colnames(df5[,1:112]),corOPLSDA_df)
colnames(corOPLSDA_df2) <- c("Lipoprotein", "OPLSDA_correlation")
OPLSDAsig <- corOPLSDA_df2 %>%
  filter(OPLSDA_correlation<(-0.5)|OPLSDA_correlation>0.5)
#compare pls-da and opls-da loadings and scores since opls-da doesn't work with the full set in df3, only pls-da does
proof<-as.data.frame((cbind(roplsOPLSDAscores$p1, roplsPLSDAscores$p1))) 


ggplot(data=proof, aes(roplsOPLSDAscores$p1, roplsPLSDAscores$p1)) + geom_point()+stat_summary(fun.data=mean_cl_normal) + 
  geom_smooth(method='lm', formula= y~x) + theme_minimal()+ggtitle("ropls Scores") + labs(x="oplsda", y="plsda")
cor.test(roplsOPLSDAscores$p1, roplsPLSDAscores$p1, method="pearson")

proof<-as.data.frame((cbind(roplsOPLSDAloadings$p1, roplsPLSDAloadings$p1)))
ggplot(data=proof, aes(roplsOPLSDAloadings$p1, roplsPLSDAloadings$p1)) + geom_point()+stat_summary(fun.data=mean_cl_normal) + 
  geom_smooth(method='lm', formula= y~x) + theme_minimal()+ggtitle("ropls Loadings") + labs(x="oplsda", y="plsda")
cor.test(roplsOPLSDAloadings$p1, roplsPLSDAloadings$p1, method="pearson")
#comparison of PLS-DA, OPLS-DA and Univariate p-values 

comp<-full_join(UnivariateSignificant,PLSDAsig,by=c("Lipoprotein"))
comp<-full_join(comp, OPLSDAsig,by=c("Lipoprotein"))
# test<-full_join(test, sigRGCCA, by=c("Lipoprotein"))


V=list(UV=UnivariateSignificant$Lipoprotein, PLSDA=PLSDAsig$Lipoprotein, OPLSDA=OPLSDAsig$Lipoprotein)
ggVennDiagram(V, color = "black", lwd = 0.8, lty = 1) + 
  scale_fill_gradient(low = "#F4FAFE", high = "#4981BF")
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
